A numerical study of the solution of X-Mode equations around the hybrid resonance

### 2015-06-10 report_clean_final/TEX/timedomain.tex
addede bruno's modifications

\section{Time-Dependent Problem Discretization}
\arev{
We develop hereafter the discretization of the time-dependent one-dimensional model (\ref{eq:main_model}), for the particular case $\theta=0$. 
We first show how this system can be discretized thanks to a Yee scheme suggested in \cite{stable_yee_plasma_current}, then a scheme based on semi-lagrangian discretization is employed.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A staggered scheme}
\label{sec:staggered_discretization}
%In this section, we develop a discretization of the time-dependent problem~\eqref{eq:main_model} in one dimension. %In 1D, the derivatives with respect to $y$ vanish, and we are left with the following problem, 
% \begin{align}
% \epsilon_0\partial_t E_{x}=-eN_e u_x, \nonumber\\
% \epsilon_0\partial_t E_{y}+\partial_x H_z=-eN_e u_y,\nonumber\\
% \mu_0\partial_t H_z+\partial_x E_y-\partial_y E_x=0, \label{eq:tdpb}\\
% m_e\partial_t u_x=eE_x+eu_yB_0-\nu m_e u_x,\nonumber\\
% m_e\partial_t u_y=eE_y-eu_xB_0-\nu m_e u_y.\nonumber
% \end{align}
We compute a numerical approximation of~\eqref{eq:main_model} \mrev{with $\theta=0$} on a cartesian grid using the Yee scheme, 
as suggested in \cite{stable_yee_plasma_current}. Let $n$ and $\Delta t$ denote the time step index and the time step, as usual the discrete time is $t_n = n \Delta t$. 
It is standard (see \cite{Xu_2006,stable_yee_plasma_current }) to discretize the electrical field on the integer time steps, 
whereas the field $H_z$ on half time steps, $t_{n+1/2}$. The time discretization of the system reads:
 \bealn
 \left\{\begin{array}{l}
 \epsilon_0\frac{ E_{x}^{n+1}-E_x^{n}}{\Delta t}=-eN_e \frac{u_x^{n}+u_x^{n+1}}{2}, \nonumber\\
 \epsilon_0\frac{ E_{y}^{n+1}-E_y^{n}}{\Delta t}+\partial_x H_z^{n+1/2}=-eN_e \frac{u_y^{n}+u_{y}^{n+1}}{2},\nonumber\\
 \mu_0\frac{ H_z^{n+1/2}-H_z^{n-1/2}}{\Delta t}+\partial_x E_y^{n}-\partial_y E_x^{n}=0, \label{eq:tdiscretizedpb}\\
 m_e\frac{ u_{x}^{n+1}-u_x^{n}}{\Delta t}=e\frac{E_x^{n+1}+E_x^{n}}{2}+e\frac{u_y^{n+1}+u_{y}^{n}}{2}B_0-\nu m_e \frac{u_x^{n+1}+u_x^{n}}{2},\nonumber\\
 m_e\frac{u_{y}^{n+1}-u_y^{n}}{\Delta t}=e\frac{E_y^{n+1}+E_y^{n}}{2}-e\frac{u_x^{n+1}+u_x^{n}}{2}B_0-\nu m_e \frac{u_y^{n+1}+u_y^{n}}{2}.\nonumber
\end{array}\right.
 \eealn
In Yee scheme, the different components of the vector fields 
are defined on staggered points. 
We summarize this discretization on Fig.~\ref{schemepos} where the positions 
of \urev{$E_x$}, $u_x$ (black dots), $E_y$, $u_y$ (blue dots) and $H_z$ (red dots) 
on the discretized grid in space and time are displayed. The fully discretized system then reads

@@

The scheme described hereafter is mentioned since it is based on some well-known 
numerical methods in plasma physics, the so-called semi-lagrangian schemes widely used for the numerical approximation of transport equations.
It seems to be an original use of semi-lagrangian schemes 
and is mentioned as possible other option. 
We restrict the presentation to the minimum and leave the details  of the construction to the alert reader.

@@

\begin{remark}
	The structure of this algorithm is very similar to the one developed in \cite{compfluids} and can be adapted and studied numerically
	in any dimension on a Cartesian grid and for more complex right-hand sides.
\end{remark}

>>>>

\section{Time-Dependent Problem Discretization}
\arev{
We develop hereafter the discretization of the time-dependent one-dimensional model (\ref{eq:main_model}), for the particular case $\theta=0$. 
We first show how this system can be discretized thanks to a Yee scheme suggested in \cite{stable_yee_plasma_current}, then a scheme based on semi-lagrangian discretization is employed.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A staggered scheme}
\label{sec:staggered_discretization}
%In this section, we develop a discretization of the time-dependent problem~\eqref{eq:main_model} in one dimension. %In 1D, the derivatives with respect to $y$ vanish, and we are left with the following problem, 
% \begin{align}
% \epsilon_0\partial_t E_{x}=-eN_e u_x, \nonumber\\
% \epsilon_0\partial_t E_{y}+\partial_x H_z=-eN_e u_y,\nonumber\\
% \mu_0\partial_t H_z+\partial_x E_y-\partial_y E_x=0, \label{eq:tdpb}\\
% m_e\partial_t u_x=eE_x+eu_yB_0-\nu m_e u_x,\nonumber\\
% m_e\partial_t u_y=eE_y-eu_xB_0-\nu m_e u_y.\nonumber
% \end{align}
We compute a numerical approximation of~\eqref{eq:main_model} \mrev{with $\theta=0$} on a cartesian grid using the Yee scheme, 
as suggested in \cite{stable_yee_plasma_current}. Let $n$ and $\Delta t$ denote the time step index and the time step, as usual the discrete time is $t_n = n \Delta t$. 
It is standard (see \cite{Xu_2006,stable_yee_plasma_current }) to discretize the electrical field on the integer time steps, 
whereas the field $H_z$ on half time steps, $t_{n+1/2}$. The time discretization of the system reads:
 \bealn
 \left\{\begin{array}{l}
 \epsilon_0\frac{ E_{x}^{n+1}-E_x^{n}}{\Delta t}=-eN_e \frac{u_x^{n+1}+u_x^{n}}{2}, \nonumber\\
 \epsilon_0\frac{ E_{y}^{n+1}-E_y^{n}}{\Delta t}+\partial_x H_z^{n+1/2}=-eN_e \frac{u_y^{n+1}+u_{y}^{n}}{2},\nonumber\\
 \mu_0\frac{ H_z^{n+1/2}-H_z^{n-1/2}}{\Delta t}+\partial_x E_y^{n}-\partial_y E_x^{n}=0, \label{eq:tdiscretizedpb}\\
 m_e\frac{ u_{x}^{n+1}-u_x^{n}}{\Delta t}=e\frac{E_x^{n+1}+E_x^{n}}{2}+e\frac{u_y^{n+1}+u_{y}^{n}}{2}B_0-\nu m_e \frac{u_x^{n+1}+u_x^{n}}{2},\nonumber\\
 m_e\frac{u_{y}^{n+1}-u_y^{n}}{\Delta t}=e\frac{E_y^{n+1}+E_y^{n}}{2}-e\frac{u_x^{n+1}+u_x^{n}}{2}B_0-\nu m_e \frac{u_y^{n+1}+u_y^{n}}{2}.\nonumber
\end{array}\right.
 \eealn
In Yee scheme, the different components of the vector fields 
are defined on staggered points. 
We summarize this discretization on Fig.~\ref{schemepos} where the positions 
of \urev{$E_x$}, $u_x$ (black dots), $E_y$, $u_y$ (blue dots) and $H_z$ (red dots) 
on the discretized grid in space and time are displayed. The fully discretized system then reads

@@

The scheme described hereafter is mentioned since it is based on some well-known 
numerical methods in plasma physics, the so-called semi-lagrangian schemes widely used for the numerical approximation of transport equations.
It seems to be an original use of semi-lagrangian schemes 
and is mentioned as possible other option. 
We restrict the presentation to the minimum and leave the details  of the construction to the alert reader.
Concerning the   vast literature on the use of semi-lagrangian schemes
in numerical plasma physics, we refer the reader to \cite{feix,guo,umeda,fifi,soso} and references therein.
	The structure of the following  algorithm is very similar to the one developed in \cite{compfluids} and can be adapted and studied numerically
	in any dimension on a Cartesian grid and for more complex right-hand sides.
	So mathematical properties of the  first step of the following semi-lagrangian scheme have been described in  \cite{cedar}.

@@

### 2015-06-10 report_clean_merged/numerics.tex
addede bruno's modifications

\item The form of the exact solution to the problem (\ref{vf1dcase}), see \cite{Despres_2014} and \cite[Proposition 1]{singular_solutions}
 E_x^{\nu}(x)=f_{\nu}(x)\log\rho_{\nu}(x-X_{\nu})+\frac{g_{\nu}(x)}{\rho(x-X_{\nu})}+S_{1}^{\nu}(x),
where $\rho_{\nu}$ is a smooth function s.t. $\rho_{\nu}(0)=0,$ with the derivative $\rho_{\nu}'(0)=1$, 
the point $X_{\nu}\approx i\nu c+O(\nu^2)$, for some $c\in\mathbb{R}_{*}$; $f_{\nu},\; g_{\nu}, \; S_{1}^{\nu}(x)$ are smooth functions, 
which, together with their derivatives, are bounded independently of $\nu$. 
\item \mrev{An explicit bound on $\|\left(E_{y}^{\nu}\right)''\|_{L^{2}(\Omega)}$ is available,  
based on the results of \cite{singular_solutions}. First of all, we
will assume that the main contribution to $\|\left(E_{y}^{\nu}\right)''\|_{L^{2}(\Omega)}$ comes from the behaviour in the vicinity of the resonance, where 
the following expansion is valid \cite[Proposition 1]{singular_solutions}
and $p_{\nu}(x),\; S_{2}^{\nu}(x)$ are smooth functions, which, together with their derivatives, are bounded independently of $\nu$.}

@@

@@

The setting of the numerical results of figure \ref{fig:vasl}
is the following. We use a 7th order semi-lagrangian scheme, and the CFL is $\nu=0.5$.
The \arev{peak} is the hybrid resonance, visible on $E_x$ and $u_x$.
The total energy $\mathcal E(t)$ is the physical energy. 
The comparison with figures \ref{fig:dependence} and \ref{fig:resonance_nus_ey_x} shows that the 
singular nature of the \arev{resonance} is captured by this new scheme without any doubt.

\begin{figure}{h}
	\begin{center}
		\begin{tabular}{ccc}
			\includegraphics[angle=-90,width=5.cm]{pics_semilagrange/Ey_moit.eps}
			&
			\includegraphics[angle=-90,width=5.cm]{pics_semilagrange/Hz_moit.eps} &
			\includegraphics[angle=-90,width=5.cm]{pics_semilagrange/Ex_moit.eps}
			\\
			$E_y$ & $H_z$  & $E_x^\nu$ \\
			%\hline
			\hline
			\includegraphics[angle=-90,width=5.cm]{pics_semilagrange/uy_moit.eps}
			&
			\includegraphics[angle=-90,width=5.cm]{pics_semilagrange/normes_moit.eps}
			&
			\includegraphics[angle=-90,width=5.cm]{pics_semilagrange/ux_moit.eps}
			\\      
			$u_y$ & $t\mapsto \mathcal E(t)$ & $u_x^\nu$  \\

>>>>

@@

\item \mrev{An explicit bound on $\|\left(E_{x}^{\nu}\right)''\|_{L^{2}(\Omega)}$ from \cite{Despres_2014} and \cite[Proposition 1]{singular_solutions}. 
First of all, we 
will assume that the main contribution to $\|\left(E_{x}^{\nu}\right)''\|_{L^{2}(\Omega)}$ comes from the behaviour in the vicinity of the resonance, where, 
for sufficiently small $\nu$, the following expansion is valid \cite[Proposition 1]{singular_solutions}
 E_x^{\nu}(x)=\tilde{E}_x^{\nu}+S_{1}^{\nu}(x),\\
 \tilde{E}_x^{\nu}=f_{\nu}(x)\log\rho_{\nu}(x-X_{\nu})+\frac{g_{\nu}(x)}{\rho(x-X_{\nu})},
where $f_{\nu},\; g_{\nu}, \; S_{1}^{\nu}(x)$ are smooth functions, 
which, together with their derivatives, are bounded independently of $\nu$ for all sufficiently small values of $\nu$; 
$\rho_{\nu}$ is a smooth function s.t. $\rho_{\nu}(0)=0,$ with the derivative $\rho_{\nu}'(0)=1$, 
the point $X_{\nu}\approx i\nu c+O(\nu^2)$, for some $c\in\mathbb{R}_{*}$. 
}
\item \mrev{An explicit bound on $\|\left(E_{y}^{\nu}\right)''\|_{L^{2}(\Omega)}$ from \cite{singular_solutions}. As before, we 
will assume that the main contribution to $\|\left(E_{y}^{\nu}\right)''\|_{L^{2}(\Omega)}$ comes from the behaviour in the vicinity of the resonance, where,
for sufficiently small $\nu$, it holds \cite[Proposition 1]{singular_solutions}
and $p_{\nu}(x),\; S_{2}^{\nu}(x)$ are smooth functions, which, together with their derivatives, are bounded independently of $\nu$ 
for all sufficiently small values of $\nu$.}

@@

The setting of the numerical results of Figure \ref{fig:vasl}
is the following. We use a 7th order semi-lagrangian scheme, and the CFL is $\nu=0.5$.
The \arev{peak} is the hybrid resonance, visible on $E_x$ and $u_x$.
The total energy $\mathcal E(t)$ is the physical energy. 
The comparison with Figures \ref{fig:dependence} and \ref{fig:resonance_nus_ey_x} shows that the 
singular nature of the \arev{resonance} is captured by this new scheme without any doubt.

\urev{Notice that in this case we shifted the computational domain so that $L=0$ and the point where the resonance occurs is $x_h=20$.}

\begin{figure}[h]
	\begin{center}
		\begin{tabular}{ccc}
			\includegraphics[angle=-90,width=5.cm]{pics_semilagrange/Ey_moit.eps}
			&
			\includegraphics[angle=-90,width=5.cm]{pics_semilagrange/Hz_moit.eps} &
			\includegraphics[angle=-90,width=5.cm]{pics_semilagrange/Ex_moit.eps}
			\\
			$E_y$ & $H_z$  & $E_x^\nu$ \\
			%\hline
			%\hline
			\includegraphics[angle=-90,width=5.cm]{pics_semilagrange/uy_moit.eps}
			&
			\includegraphics[angle=-90,width=5.cm]{pics_semilagrange/normes_moit.eps}
			&
			\includegraphics[angle=-90,width=5.cm]{pics_semilagrange/ux_moit.eps}
			\\      
			$u_y$ & $t\mapsto \mathcal E(t)$ & $u_x^\nu$  \\
